Effects of Different Pain Relievers and Dosages on Cognitive Retention

### 2025-04-09 paper/paper.qmd
Power/Effect Size
Limitations

Our study is overall focused on memory, which may have links to one’s education and IQ level, as they are all related to mental function and cognition. In the future, we could test everyone for IQ scores and their highest level of education – these two nuisance variables can be taken care of by using the four principles of experimental design.
Additionally, we used 30 observations for each treatment group which may not be large enough. We can increase this number for future studies. To add on, not all participants were in pain, so the efficacy of the improvement of pain killers may be biased and using a virtual population may not ensure whether this aspect could be captured.
Lastly, cognitive retention was measured immediately post-treatment – the long term impacts are not yet explored. To improve our study, we can try to consider individuals with real pain symptoms before the study to ensure a reduction of bias, and provide a realistic test setting. Additionally, carrying the experiment out in the long-term would help with understanding the effects of painkillers and dosages on memory.

There’s also a chance that the participants’ expectations influenced their results if they knew which treatment they were getting (we cannot expand on this as it was a virtual experiment). In the future, bias could be minimized through having a double-blind design in which neither the participants nor the researchers know who’s getting what. We only tested low and high dosages, so we might be missing more subtle effects that show up with intermediate doses. So if we were to include a wider range of dosages in future studies, it could give a fuller picture. Another issue is that our sample might not fully represent the general population since we selected participants from specific clusters of locations -- not all locations were covered equally. This means our findings might not apply to everyone. Next time, we could choose more diverse and evenly spread-out clusters to make the results more general. Lastly, despite our efforts to control for confounding factors, there are still some variables (like sleep quality, stress levels, or baseline cognitive function) that could affect memory performance. Being more thorough in accounting for these factors could improve the accuracy of the results.

>>>>

Our study is overall focused on memory, which may have links to one’s education and IQ level, as they are all related to mental function and cognition. In the future, we could test everyone for IQ scores and their highest level of education – these two nuisance variables can be taken care of by using the four principles of experimental design.

Additionally, we used 30 observations for each treatment group, which may not be large enough. We can increase this number for future studies. In fact, our post-hoc power analysis showed that the statistical power for all three effects were quite low (ranging between 20% and 38%) with the effect sizes falling in the small-medium range. This suggests that the insignificance of our findings may have been due to limited power instead of the absence of a true effect. Future studies could prevent this issue by increasing the sample size or using stronger manipulations in order to improve the likelihood of detecting meaningful effects. To add on, not all participants were in pain, so the efficacy of the improvement of pain killers may be biased and using a virtual population may not ensure whether this aspect could be captured.

Lastly, cognitive retention was measured immediately post-treatment – the long term impacts are not yet explored. To improve our study, we can try to consider individuals with real pain symptoms before the study to ensure a reduction of bias, and provide a realistic test setting. Additionally, carrying the experiment out in the long-term would help with understanding the effects of painkillers and dosages on memory.

There’s also a chance that the participants’ expectations influenced their results if they knew which treatment they were getting (we cannot expand on this as it was a virtual experiment). In the future, bias could be minimized through having a double-blind design in which neither the participants nor the researchers know who’s getting what. We only tested low and high dosages, so we might be missing more subtle effects that show up with intermediate doses. So if we were to include a wider range of dosages in future studies, it could give a fuller picture.

Another issue is that our sample might not fully represent the general population since we selected participants from specific clusters of locations -- not all locations were covered equally. This means our findings might not apply to everyone. Next time, we could choose more diverse and evenly spread-out clusters to make the results more general.

Lastly, despite our efforts to control for confounding factors, there are still some variables (like sleep quality, stress levels, or baseline cognitive function) that could affect memory performance. Being more thorough in accounting for these factors could improve the accuracy of the results.
